Guidelines for Submissions to ISCA 2013

### 2012-11-18 paper/isca40format/stuff/Threads.tex
stuff

The behavior of each node can be best described as a system of suspendable local
threads, with a scheduler local to that node.  A thread can be in one of the two
states: \emph{executing} or \emph{suspended}. Once a thread finishes executing,
the scheduler does one of the following: a) picks one of suspended threads which
has become ready to execute due to changed conditions or b) creates a new thread
to handle an incoming message. An executing thread can read the local state
of the node, send messages to output channels, receive messages from input
channels or spawn more threads. The input and output channels are blocking, so
if a thread tries to send to a full output channel or receive from an empty
input channel, the thread would not be able to proceed. The thread can also
fail to proceed due to some condition in the local state of the node. In
all these cases, the thread gets suspended. It remains suspended till the
required conditions are met (for instance, the output channel becomes free, the
input channel receives a pertinent message, or other local state changes),
Once the required conditions for a suspended thread are met the scheduler may
pick it for execution. The execution is resumed from the exact point where it
was suspended. Once the thread finishes its execution, all the book-keeping
resources associated with that thread is freed up and reclaimed.

Incoming messages remain in the channel until they are explicitly removed by
the thread dealing with that message. If the scheduler sees an incoming message
that requires a new thread to be created for handling, and if the required
book-keeping resources are available, then a thread is created.
Different threads require different amounts of book-keeping resources. If the
required book-keeping resources are not available, the scheduler might pick
another suspended thread which is ready to execute, or if that's not possible,
spin-waits.  The threads and incoming messages have a memory address associated
with them. For an incoming message that does not require a new thread to be
created, its address is used by the scheduler to decide if a suspended thread
is ready to execute. Similarly, the address associated with a thread that has
just finished executing is used to determine if the local state of the node
associated with that address has changed so that a suspended thread associated
with that address can resume execution.

@@

Earlier, each thread is started,, because of an incoming message, with a
handler \textbf{procedure} particular to the type of the message. A procedure
groups together actions which include state updates ($s \gets value$),
\textbf{send} messages into output channels and \textbf{pop} messages from
input channels. \textbf{send} and \textbf{pop} are blocking calls, as mentioned
earlier. Procedures can \textbf{call} other procedures but they can not be
recursive.

>>>>

The behavior of each node can be best described as a system of suspendable local
threads, with a scheduler local to that node. A thread can be in one of the two
states: \emph{executing} or \emph{suspended}. At a given time, either one
thread or the scheduler will be executing. When the scheduler is executing,
it does one of the following: a) picks one of suspended threads
which has become ready to execute due to changed conditions or b) creates a new
thread to handle an incoming message. An executing thread can read the local
state of the node, send messages to output channels, receive messages from
input channels, \etc% or spawn more threads . The input and output channels are
blocking, so if a thread tries to send to a full output channel or receive from
an empty input channel, the thread would not be able to proceed. The thread can
also fail to proceed due to some condition in the local state of the node. In
all these cases, the thread gets suspended, giving back control to the
scheduler. It remains suspended till the required conditions are met (for
instance, the output channel becomes free, the input channel receives a
pertinent message, or other local state changes), Once the required conditions
for a suspended thread are met the scheduler may pick it for execution. The
execution is resumed from the exact point where it was suspended. Once the
thread finishes its execution, all the book-keeping resources associated with
that thread is freed up and reclaimed.

Incoming messages remain in the channel until they are explicitly removed by
the thread dealing with that message. If the scheduler sees an incoming message
that requires a new thread to be created for handling, and if the required
book-keeping resources are available, then a thread is created.
%Different threads require different amounts of book-keeping resources.
If the
required book-keeping resources are not available, the scheduler might pick
another suspended thread which is ready to execute, or if that's not possible,
spin-waits.  The threads and incoming messages have a memory address associated
with them. For an incoming message that does not require a new thread to be
created, its address is used by the scheduler to decide if a suspended thread
is ready to execute. Similarly, the address associated with a thread that has
just finished executing is used to determine if the local state of the node
associated with that address has changed so that a suspended thread associated
with that address can resume execution.

@@

\begin{figure}
\begin{tabularx}{\linewidth}{|l|X|}
\hline
\send{} $m$ & Send a message $m$ to the node's output channel. This action
blocks till the channel has space to send.\\
\receive{} $m$ & Receive and remove a message $m$ from the node's input channel.
This action blocks till a message is received.\\
$x \gets value$ & Local state $x$ of the node gets updates to $value$.\\
\call{} $f(args)$ & Execute \textbf{procedure} $f$ with $args$ arguments.\\
\start{} $f(args)$ & Start a new thread that executes \textbf{procedure} $f$
with $args$ arguments. This action is performed only by the scheduler.\\
\resume{} $t$ & Resume thread $t$ from where it stopped execution. This action
is performed only by the scheduler\\
\hline
\end{tabularx}
\caption{Actions used for describing a node's behavior}
\label{actions}
\end{figure}

When a new message arrives in an input channel that requires creation of a
thread, the scheduler first removes the message from the channel and then
creates a thread that starts executing the \procedure{} specific to the type of
the incoming message. Figure \ref{actions} show the notations we use to
describe the actions performed by an executing thread. Procedures can \call{}
other procedures but they can not be recursive. We usually denote the execution
of actions \receive{} and \send{} using their verb forms, for instance,
,``receiving'' or ``sending''. We also denote the thread created by the
scheduler to handle an incoming message as the thread ``handling'' that
message.
